---
title: "Explore Admin, Dist, C-flux df"
author: "Sam Woodman"
editor_options:
  chunk_output_type: console
output: html_document
---

## Description: 

Using the dataframe created from the stacked admin, dist, C-flux rasters we 
can explore some initial plots and sumamries of the data. Questions and findings
here will provide a starting off point for future analyses. The ultimate aim of
the project is to determine if land management and disturbance interact to 
increase or decrease C-flux across the landscape.

### Packages 

```{r}
library(tidyverse)
library(here)
library(knitr)
library(ggridges)
```


### Load data

This data represents 250m pixels of total C-flux from 2001 to 2019, number of 
disturbance events that have occurred in each pixel for that time period, and 
the land management type. 

```{r}
cad_df <- read_csv(here("data/processed/cflux_admin_dist_df.csv"))
```


#### How many pixels of each land zone are within the data?

```{r}
cad_df %>% 
  group_by(zone) %>% 
  tally(name = "n_pixels") %>% 
  mutate(area = n_pixels*(250*250)) %>% 
  arrange(desc(area)) %>% 
  kable()

zone_to_remove <- c("National Park", "Wilderness Area", 
                    "National Marine Conservation Area", 
                    "National Wildlife Area")
```

Four land zones have fewer than 100 pixels. Given the magnitude of the other 
groups these four should likely be removed.

#### What is the distribution of disturbance counts across zones

```{r}
cad_df %>% 
  filter(!zone %in% zone_to_remove) %>% 
  #mutate(zone = str_replace(zone, "Recommended ", "")) %>% 
  group_by(zone, count) %>% 
  tally(name = "n_pixels") %>% 
  mutate(freq = n_pixels/sum(n_pixels)) %>% 
  ggplot(aes(x = count, y = freq)) + 
  geom_col() + 
  labs(x = "Number of disturbances", 
       y = "Proportion of events")  +
  facet_wrap(~zone) + 
  theme_bw()

cad_df %>% 
  filter(!zone %in% zone_to_remove) %>% 
  mutate(zone = str_replace(zone, "Recommended ", "")) %>% 
  ggplot(aes(x = count, y = zone)) + 
  geom_density_ridges(rel_min_height = 0.01) + 
  theme_bw()
```

#### What proportion of pixels experience disturbance?

```{r}
cad_df %>% 
  filter(!zone %in% zone_to_remove) %>% 
  mutate(zone = str_replace(zone, "Recommended ", "")) %>% 
  group_by(zone, count) %>% 
  tally(name = "n_pixels") %>% 
  mutate(dist_binary = if_else(count >= 1, "Disturbed", "Non-disturbed")) %>% 
  group_by(zone, dist_binary) %>% 
  summarise(n_pixels = sum(n_pixels)) %>% 
  mutate(freq = n_pixels/sum(n_pixels)) %>% 
  filter(dist_binary == "Disturbed") %>% 
  ggplot(aes(x = freq, y = reorder(zone, freq))) + 
  geom_col() + 
  labs(x = "Proportion of pixels disturbed", 
       y = "") + 
  theme_bw()
```


#### How often do disturabnces occur across each land zone?

```{r}
## plot mean ± se for number of disturbances by zone
cad_df %>% 
  filter(!zone %in% zone_to_remove) %>% 
  group_by(zone) %>% 
  summarise(mean_dist = mean(count),
            sd_dist = sd(count),
            n_dist = n(),
            se_dist = sd_dist/sqrt(n_dist),
            quant = quantile(count, c(0.025, 0.5, 0.975)), 
            q = c(0.025, 0.5, 0.975)) %>% 
  pivot_wider(names_from = q, values_from = quant, names_prefix = "q_") %>% 
  ggplot(aes(x = mean_dist, y = reorder(zone, mean_dist), size = n_dist)) + 
  geom_point() + 
  geom_errorbar(aes(xmin = mean_dist - se_dist, xmax = mean_dist + se_dist), 
                size = 0.5) +
  ylab("") + 
  xlab("Mean (± SE) number of disturbances") + 
  labs(size = "Number of\npixels") + 
  theme_bw()


## boxplot of number of disturbances by zone
cad_df %>% 
  filter(!zone %in% zone_to_remove) %>% 
  group_by(zone) %>% 
  ggplot(aes(x = count, y = reorder(zone, count))) + 
  geom_boxplot() + 
  ylab("") + 
  xlab("Mean (± SE) number of disturbances") + 
  theme_bw()

```

Far north protected areas appear to expereince the most disturbances while 
provincial parks experience the fewest. However, as illustrated by the point size, 
there is a clear imbalance in the data resulting in some zone having much narrower
error bars. 


#### Does combining recommended and existing zone change the data substantially?

```{r}
cad_df %>% 
  filter(!zone %in% zone_to_remove) %>% 
  mutate(zone = str_replace(zone, "Recommended ", "")) %>% 
  group_by(zone) %>% 
  tally(name = "n_pixels") %>% 
  mutate(area = n_pixels*(250*250)) %>% 
  arrange(desc(area)) %>% 
  kable()

cad_df %>% 
  filter(!zone %in% zone_to_remove) %>% 
  mutate(zone = str_replace(zone, "Recommended ", "")) %>% 
  group_by(zone) %>% 
  summarise(mean_dist = mean(count),
            sd_dist = sd(count),
            n_dist = n(),
            se_dist = sd_dist/sqrt(n_dist),
            quant = quantile(count, c(0.025, 0.5, 0.975)), 
            q = c(0.025, 0.5, 0.975)) %>% 
  pivot_wider(names_from = q, values_from = quant, names_prefix = "q_") %>% 
  ggplot(aes(x = mean_dist, y = reorder(zone, mean_dist), size = n_dist)) + 
  geom_point() + 
  geom_errorbar(aes(xmin = mean_dist - se_dist, xmax = mean_dist + se_dist), 
                size = 0.5) +
  ylab("") + 
  xlab("Mean (± SE) number of disturbances") + 
  labs(size = "Number of\npixels") + 
  theme_bw()
```

Combining Recommended Provincial Parks with Provincial Parks and Recommended 
Conservation Reserves with Conservation Reserves naturally increase the pixel 
count for each group. For summary of number of disturbances the only noticeable
change is that Recommended Provincial Parks, which experienced more disturbances
than General Use Area, were incorporated into Provincial Parks. 




