---
title: "Explore C-flux Results"
author: "Sam Woodman"
editor_options:
  chunk_output_type: console
output: html_document
theme: lumen
---

```{r include = FALSE}
knitr::opts_chunk$set(echo=FALSE)
```


## Description: 

Using the dataframe created from per hectare C-flux and disturbance history 
(2001 - 2019) for land administration zones in Ontario we explore possible 
questions to include in a future manuscript. The ultimate aim of the project is 
to determine if land management and disturbance interact to increase or decrease 
C-flux across the landscape.

### Packages 

```{r packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
library(knitr)
library(ggridges)
library(ggsci)
library(spaMM)
library(modelbased)
```


### Load data

This data represents mean megagrams of CO~2~ emissions per ha from 2001 to 2019, 
for land management zones in Ontario. The mean proportion of each management 
zone was also calculated across the period of C-flux data (2001 - 2019).

```{r load, message=FALSE, warning=FALSE}
cad_poly_df <- read_csv(here("data/processed/admin_cflux_dist_poly_ha_df.csv")) 
cad_poly_df$zone <- relevel(as.factor(cad_poly_df$zone), "General Use Area")
skimr::skim(cad_poly_df)

admin_dist_df_year <- read_csv(here("data/processed/dist_admin_extract_year.csv")) %>% 
  mutate(zone = as.character(zone),
         zone = if_else(zone == "Indian Reserve", "Indigenous Reserve", zone),
         zone = as.factor(zone))
admin_dist_df_year$zone <- relevel(as.factor(admin_dist_df_year$zone), "General Use Area")

```

#### What does each land management zone mean in terms of regualtion and protection?

* General Use Area
  + 
* Enhanced Management Area
  + Remote access enhanced management areas are intended to maintain the remote character of selected areas. Typically, these are relatively large areas which provide the public and tourism operators with high-quality remote recreational experiences including hinting, fishing, canoeing, and camping. Given the large size, remoteness, and relative absence of roads, these areas will play a significant role in protecting wilderness values outside the parks and protected areas system.
* Indigenous Reserves [1](https://www.sac-isc.gc.ca/eng/1587563382582/1587563402582)
  + [Provincial law requires consultation with First Nation and Metis people 
  during environmental assessment](https://www.ontario.ca/page/environmental-assessments-consulting-indigenous-communities)
* Provincial Parks [1](http://bmfci.ca/phocadownloadpap/Bancroft-HCV-assessment-Vers1-2012Sep.pdf) [2](https://www.iucn.org/downloads/ontario.pdf)
  + Regulated and forestry activities not allowed
  + Objectives set by Ontario government
    + To permanently protect representative ecosystems, biodiversity and provincially significant elements of Ontario’s natural and cultural heritage, and to manage these areas to ensure that ecological integrity is maintained.
    + To provide opportunities for ecologically sustainable outdoor recreation opportunities and encourage associated economic benefits.
    + To provide opportunities for residents of Ontario and visitors to increase their knowledge and appreciation of Ontario’s natural and cultural heritage.
    + To facilitate scientific research and provide points of reference to support the monitoring of ecological change on the broader landscape.30
* Conservation Reserves [1](http://bmfci.ca/phocadownloadpap/Bancroft-HCV-assessment-Vers1-2012Sep.pdf) [2](https://www.iucn.org/downloads/ontario.pdf)
  + Regulated and forestry activites not allowed
  + Objectives set by Ontario government
    + To permanently protect representative ecosystems, biodiversity and provincially significant elements of Ontario’s natural and cultural heritage, and to manage these areas to ensure that ecological integrity is maintained.
    + To provide opportunities for ecologically sustainable land uses, including traditional outdoor heritage activities and associated economic benefits.
    + To facilitate scientific research and provide points of reference to support the monitoring of ecological change on the broader landscape.36


#### Are administration zones consistent in size

```{r size_analysis}
cad_poly_df %>% 
  group_by(zone) %>% 
  summarise(mean_area = mean(area_ha, na.rm = T),
            sd_area = sd(area_ha),
            n_area = n(),
            se_area = sd_area/sqrt(n_area)) %>% 
  ggplot(aes(x = mean_area, y = zone)) + 
  geom_col() + 
  geom_errorbar(aes(xmin = mean_area - se_area, 
                    xmax = mean_area + se_area)) +
  scale_x_log10(expand = c(0,0),
                limits = c(1, 1000000)) + 
  xlab("Log mean (±SE) area (ha)") + 
  ylab("") + 
  theme_bw() + 
  theme(text = element_text(size=12))

cad_poly_df %>% 
  ggplot(aes(x = area_ha, y = zone)) + 
  geom_density_ridges(rel_min_height = 0.01) +
  scale_x_log10() +
  xlab("Log area (ha)") + 
  ylab("") + 
  theme_bw()

summary(lm(log(area_ha) ~ zone, data = cad_poly_df))
estimate_contrasts(lm(log(area_ha) ~ zone, data = cad_poly_df), 
                   transform = "response")
```

Administration zones are not equal in size across the AoU. General Use Areas are
the largest while Indigenous Reserves, Provincial Parks, and Conservation 
Reserves are the smallest. 

The difference in size between the zones suggests that a per area approach is 
needed for modelling to standardize C-fluxes.


#### Are administration zones evenly distributed across Ontario?
```{r space_analysis}
cad_poly_df %>% 
  group_by(zone) %>% 
  summarise(mean_lat = mean(lat, na.rm = T),
            sd_lat = sd(lat),
            n_lat = n(),
            se_lat = sd_lat/sqrt(n_lat),
            mean_lon = mean(lon, na.rm = T),
            sd_lon = sd(lon),
            n_lon = n(),
            se_lon = sd_lon/sqrt(n_lon)) %>% 
  ggplot(aes(x = mean_lon, y = mean_lat, colour = zone)) + 
  geom_point(size = 2) + 
  geom_errorbar(aes(ymin = mean_lat - se_lat, 
                    ymax = mean_lat + se_lat, 
                    width = 0.2)) + 
  geom_errorbarh(aes(xmin = mean_lon - se_lon, 
                     xmax = mean_lon + se_lon, 
                     height = 0.05)) + 
  xlab("Mean (±SE) longitude") + 
  ylab("Mean (±SE) latitude") + 
  labs(colour = "")  +
  scale_color_jco() + 
  theme_bw() + 
  theme(text = element_text(size=12))


summary(lm(lat ~ zone, data = cad_poly_df))
estimate_contrasts(lm(lat ~ zone, data = cad_poly_df))
summary(lm(lon ~ zone, data = cad_poly_df))
estimate_contrasts(lm(lon ~ zone, data = cad_poly_df))

```

Administration zones are not evenly spread across space. Indigenous Reserves, on
average, are located at more northern latitudes and more western longitudes than
the other zones. Conversely, Enhanced Management Areas are, on average, at more
eastern longitudes. 

Though these difference may suggest something about the location of each 
administration zone across Ontario (for example where different management 
strategies are favoured) any difference in lat/lon coordiantes will be accounted 
for in spatail models. 

#### Do administration zones experience different proportions of disturbance?

```{r prop_dist_analysis, message=FALSE, warning=FALSE}
admin_dist_df_year %>% 
  group_by(zone, Year) %>% 
  summarise(mean_prop_dist = mean(prop_dist, na.rm = T),
            sd_prop_dist = sd(prop_dist, na.rm = T),
            n_prop_dist = n(),
            se_prop_dist = sd_prop_dist/sqrt(n_prop_dist)) %>% 
  ggplot(aes(x = Year, y = mean_prop_dist)) + 
  geom_point() +
  geom_errorbar(aes(ymin = mean_prop_dist - se_prop_dist,
                    ymax = mean_prop_dist + se_prop_dist)) + 
  xlab("") + 
  ylab("Mean (±SE) proportion disturbed") +
  facet_wrap(~ zone) + 
  theme_bw() + 
  theme(text = element_text(size=12))

admin_dist_df_year %>% 
  group_by(zone, Year) %>% 
  summarise(mean_prop_dist = mean(prop_dist, na.rm = T)) %>% 
  pivot_wider(names_from = zone, values_from = mean_prop_dist) %>% 
  janitor::clean_names() %>% 
  GGally::ggpairs(., 2:6)

admin_dist_df_year %>% 
  ggplot(aes(x = prop_dist, y = zone)) + 
  geom_density_ridges(rel_min_height = 0.01) + 
  xlab("Mean proportion disturbed from 2001 to 2019") + 
  ylab("") + 
  facet_wrap(~ as.factor(Year), scales = "free_x") + 
  theme_bw()
  
summary(glm(prop_dist ~ zone * Year, family = binomial(link = "sqrt"), 
            data = admin_dist_df_year))
```

No, between the five administration zones in the study there appear to be no 
differences in the mean proportion disturbed. 

#### Do administration zones show different frequencies of disturbance

```{r n_dist_analysis}
cad_poly_df %>% 
  ggplot(aes(x = n_dist, y = zone)) + 
  geom_boxplot() +
  xlab("Number of years between 2001 and 2019 with disturbance") + 
  ylab("") +
  theme_bw() + 
  theme(text = element_text(size=12))

cad_poly_df %>% 
  group_by(zone) %>% 
  summarise(num_dist = mean(n_dist, na.rm = T),
            sd_dist = sd(n_dist, na.rm = T),
            n_dist = n(),
            se_dist = sd_dist/sqrt(n_dist)) %>% 
  ggplot(aes(x = num_dist, y = zone)) + 
  geom_col() + 
  geom_errorbar(aes(xmin = num_dist - se_dist, xmax = num_dist + se_dist)) +
  scale_x_continuous(expand = c(0,0),
                     limits = c(0, 10),
                     breaks = seq(0,10, 2)) +
  xlab("Mean (±SE) number of years between\n2001 and 2019 with disturbance") + 
  ylab("") + 
  theme_bw() + 
  theme(text = element_text(size=12))

cad_poly_df %>% 
  ggplot(aes(x = n_dist, y = zone)) + 
  geom_density_ridges(rel_min_height = 0.01) + 
  xlab("Number of years between 2001 and 2019 with disturbance") + 
  ylab("") + 
  theme_bw()

summary(glm(n_dist ~ zone, family = "quasipoisson", data = cad_poly_df))
modelbased::estimate_contrasts(glm(n_dist ~ zone, family = "poisson", 
                                   data = cad_poly_df))
```

Yes, Conservation Reserves and Indigenous Reserves experience the fewest years 
with disturbance while General Use Area has the most. These results held when 
fitting a spatial model (spaMM package, not shown here). 

#### Does disturbance alter C-flux of land administration zones?

```{r spatial_model}
spatial_mod <-
  fitme(
    cflux_mean ~ zone*sqrt(mean_prop_dist) +
      Matern(1 |lon + lat),
    data = cad_poly_df,
    family = "gaussian", method = "REML"
  ) 

summary(spatial_mod)
anova(aov(spatial_mod))

newdat <- expand_grid(zone = cad_poly_df %>% pull(zone) %>% unique(), 
                      mean_prop_dist = sqrt(seq(0,0.5,0.001)), 
                      lon = cad_poly_df %>% pull(lon) %>% mean,
                      lat = cad_poly_df %>% pull(lat) %>% mean)

newdat$pred <- as.vector(predict(spatial_mod, newdat))

newdat %>% 
  ggplot(aes(x = (mean_prop_dist)^2, y = pred, colour = zone)) + 
  geom_line(size = 1.5) + 
  ggsci::scale_color_jco() +
  xlab("Mean proportion disturbed from 2001 to 2019") + 
  ylab("Megagrams CO2 emissions/ha") + 
  labs(colour = "Land\nManagement") + 
  theme_bw() + 
  theme(text = element_text(size=12))
```

Yes, if we use General Use Area as a control then Conservation Reserves and 
Provincial Parks are less impacted by natural disturbances. CR and PP are also
the best assimilator of CO~2~ with the lowest y-intercept. Low proportions of 
disturbance appear to have a positive effect on CO~2~ assimilation though this 
trend weakens as disturbance increases. As disturbance reaches the maximum of 
the range seen in our data Enhanced Management Reserves become the best 
CO~2~ assimilators while Indigenous Reserves are the worst. 

#### Are the results different if number of years with disturbance is used 
instead of mean proportion area disturbed? 

```{r}
spatial_mod_ndist <-
  fitme(
    cflux_mean ~ zone*sqrt(n_dist) +
      Matern(1 |lon + lat),
    data = cad_poly_df,
    family = "gaussian", method = "REML"
  ) 

summary(spatial_mod_ndist)
anova(aov(spatial_mod_ndist))

newdat <- expand_grid(zone = cad_poly_df %>% pull(zone) %>% unique(), 
                      n_dist = sqrt(seq(0,19,1)), 
                      lon = cad_poly_df %>% pull(lon) %>% mean,
                      lat = cad_poly_df %>% pull(lat) %>% mean)

newdat$pred <- as.vector(predict(spatial_mod_ndist, newdat))

newdat %>% 
  ggplot(aes(x = (n_dist)^2, y = pred, colour = zone)) + 
  geom_line(size = 1.5) + 
  ggsci::scale_color_jco() +
  xlab("No. years disturbed from 2001 to 2019") + 
  ylab("Megagrams CO2 emissions/ha") + 
  labs(colour = "Land\nManagement") + 
  theme_bw() + 
  theme(text = element_text(size=12))
```


There is no longer a significant interaction between zone and disturbance history
when using number of years with disturbance. This is not surprising since number 
of years fails to capture the extent/severity of disturbance. 
